Biden Projected to Win the 2024 US Presidential Elections

### 2024-03-16 paper/paper.qmd
Final changes

Our survey data is from the Ipsos Poll [Roper #31120637]. It was collated from a web-based survey conducted on November 3 - 4, 2023 using KnowledgePanel. The survey is conducted by Ipsos and the sponsor is ABC News. This poll is based on a nationally representative probability sample of 949 adults aged 18 or older with oversamples among Black and Hispanic respondents. The poll is a collection of demographic information and political preferences.

@@

## Post-stratification Data

Our post-stratification data is from the most recent American Community Survey (ACS), the 2022 Sample. We use this to represent the voter population and adjust the weight of the survey data. The population for this sample was the entire US population, categorized into households as subgroups, and the responses of the sampled subgroups were collected. ACS 2022 dataset consists of 3,373,378 observations. Using a 1-in-100 national random sample, the ACS 2022 sample provides a thorough overview of the demographic, socioeconomic, and geographic features of the US population. Most importantly, this dataset includes people who live in both home and group quarters, offering a comprehensive view of the American population. A person becomes an entry in this dataset if they are a part of the sampled subgroups and have responded to the survey.

@@

After we created the model, we then post-stratify these proportions and to conclude who we predict to win the 2024 US Presidential Elections. Post-stratification is a statistical technique employed to refine survey estimates by aligning them with known population characteristics. It helps us ensure that our findings accurately represent the diversity of the entire population.

@@

ggplot(age_vote_for_biden, aes(x = age, y = biden_predict)) +
  geom_col(fill = "purple") +
  labs(title = "Proportions of Votes for Biden by age group",
       y = "Proportion of Votes",
       x = "Age groups")

@@

ggplot(sex_vote_for_biden, aes(x = sex, y = biden_predict)) +
  geom_col(fill = "red") +
  labs(title = "Proportions of Votes for Biden by sex",
       y = "Proportion of Votes",
       x = "Sex groups")

@@

ggplot(general_data, aes(x = as.factor(vote_biden), y = prop)) +
  geom_col(fill = "blue") +
  labs(title = "Proportions of Votes",
       y = "Proportion",
       x = "Vote") +
  scale_x_discrete(labels = c("Vote for Trump", "Vote for Biden"))

@@

For our poststratification dataset from ACS, there was no dataset available from 2023. Our poststratification data relied on the survey that was conducted in 2022, and Americans' viewpoints on their choices for political parties may have changed in a year.

@@

4. Education levels were standardized to facilitate analysis. Respondents' education information (ppeducat) was categorized into three groups: "High school", "More than high school" (including "Bachelor's degree or higher" and "Some college"), and "Less than high school". This was done to match the education level separation available in the poststratification dataset. The final column was named education.

@@

6. Gender information (ppgender) was just had a change of turning the value to lower case. The final column was named sex as just having these two criterias implies that and to maintain with the poststratification data reporting sex as well.

@@

For the poststratification dataset, the cleaning steps we took were:

@@

7. The final poststratification dataset contained the above mentioned 5 columns: age, education, income, state and sex.

@@

## Survey dataset
Here is a glimpse of the survey data set used

```{r}
#| echo: false
#| warning: false
#| message: false

head(survey_data)  %>%
  kable(digits = 2)

@@

## Poststratification dataset
Here is a glimpse of the poststratification data set used
```{r}
#| echo: false
#| warning: false
#| message: false

>>>>

Our survey data is from the Ipsos Poll [Roper #31120637]. It was collated from a web-based survey conducted on November 3 - 4, 2023 using KnowledgePanel. The survey is conducted by Ipsos and the sponsor is ABC News. This poll is based on a nationally representative probability sample of 949 adults aged 18 or older with oversamples among Black and Hispanic respondents. The poll is a collection of demographic information and political preferences [@citeIpsos].

@@

## Post-Stratification Data

Our post-stratification data is from the most recent annual American Community Survey (ACS), the 2022 sample, available through IPUMS USA [@citeACS]. We use this to represent the voter population and adjust the weight of the survey data. The population for this sample was the entire US population, categorized into households as subgroups, and the responses of the sampled subgroups were collected. ACS 2022 dataset consists of 3,373,378 observations. Using a 1-in-100 national random sample, the ACS 2022 sample provides a thorough overview of the demographic, socioeconomic, and geographic features of the US population. Most importantly, this dataset includes people who live in both home and group quarters, offering a comprehensive view of the American population. A person becomes an entry in this dataset if they are a part of the sampled subgroups and have responded to the survey.

@@

After we created the model, we then post-stratify these proportions and to conclude who we predict to win the 2024 US Presidential Elections. Post-Stratification is a statistical technique employed to refine survey estimates by aligning them with known population characteristics. It helps us ensure that our findings accurately represent the diversity of the entire population.

@@

ggplot(age_vote_for_biden, aes(x = age, y = biden_predict)) +
  geom_col(fill = "purple") +
  labs(y = "Proportion of Votes",
       x = "Age groups")

@@

ggplot(sex_vote_for_biden, aes(x = sex, y = biden_predict)) +
  geom_col(fill = "red") +
  labs(y = "Proportion of Votes",
       x = "Sex groups")

@@

ggplot(general_data, aes(x = as.factor(vote_biden), y = prop)) +
  geom_col(fill = "blue") +
  labs(y = "Proportion",
       x = "Vote") +
  scale_x_discrete(labels = c("Vote for Trump", "Vote for Biden"))

@@

For our post-stratification dataset from ACS, there was no dataset available from 2023. Our post-stratification data relied on the survey that was conducted in 2022, and Americans' viewpoints on their choices for political parties may have changed in a year.

@@

4. Education levels were standardized to facilitate analysis. Respondents' education information (ppeducat) was categorized into three groups: "High school", "More than high school" (including "Bachelor's degree or higher" and "Some college"), and "Less than high school". This was done to match the education level separation available in the post-stratification dataset. The final column was named education.

@@

6. Gender information (ppgender) was just had a change of turning the value to lower case. The final column was named sex as just having these two criterias implies that and to maintain with the post-stratification data reporting sex as well.

@@

For the post-stratification dataset, the cleaning steps we took were:

@@

7. The final post-stratification dataset contained the above mentioned 5 columns: age, education, income, state and sex.

@@

## Survey dataset
Here is a glimpse of the survey dataset used

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-survey
#| tbl-cap: Cleaned survey dataset
#| 
head(survey_data)  %>%
  kable(digits = 2)

@@

## Post-Stratification dataset
Here is a glimpse of the post-stratification data set used
```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-post-stratification
#| tbl-cap: Cleaned post-stratification dataset
